---
title: "Lab10 Covid 19 report maps"
author: "Katherine Chacon-Vargas"
date: "4/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warmings=FALSE, message=FALSE, echo=TRUE, include=FALSE}
library(tidyverse)
library(maps)
library(mapdata)
library(lubridate)
library(viridis)
library(wesanderson)
library(RColorBrewer)
```

##World COVID-19 Confirmed cases March 31,2020
```{r warnings=FALSE, message=FALSE, include=FALSE, echo=TRUE}
daily_report_03_31_2020 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-31-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  select(-Admin2, -FIPS)

a <- ggplot(daily_report_03_31_2020, aes(x = Long, y = Lat, size = Confirmed/1000)) +
    borders("world", colour = NA, fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='blue', fill='blue', alpha = 0.5) +
    labs(title = 'World COVID-19 Confirmed cases March 31, 2020',x = '', y = '',
        size="Cases (x1000))") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```

```{r a, echo=FALSE, warnings=FALSE, message=FALSE}
a
```


##COVID-19 Confirmed Cases in the US by March 31, 2020
```{r, warnings=FALSE, message=FALSE, include=FALSE, echo=TRUE}
daily_report_03_31_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-31-2020.csv")) %>% 
  select(-Admin2, -FIPS) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  filter (!Province_State %in% c("Alaska","Hawaii", "American Samoa",
                  "Puerto Rico","Northern Mariana Islands", 
                  "Virgin Islands", "Recovered", "Guam", "Grand Princess",
                  "District of Columbia", "Diamond Princess")) %>% 
  filter(Lat > 0) 

 b <- ggplot(daily_report_03_31_2020, aes(x = Long, y = Lat, size = Confirmed/1000)) +
    borders("state", colour = "black", fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='blue', fill='blue', alpha = 0.5) +
    labs(title = 'COVID-19 Confirmed Cases in the US by March 31, 2020', x = '', y = '',
        size="Cases (x1000))") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```
```{r b, echo=FALSE, warnings=FALSE, message=FALSE}
b
```


##COVID-19 Confirmed Cases in the US April 12, 2020
```{r, warnings=FALSE, message=FALSE, include=FALSE, echo=TRUE}
mybreaks <- c(1, 100, 1000, 10000, 10000)
daily_report_04_12_2020 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-12-2020.csv"))%>% 
  select(-Admin2, -FIPS) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  filter (!Province_State %in% c("Alaska","Hawaii", "American Samoa",
                  "Puerto Rico","Northern Mariana Islands", 
                  "Virgin Islands", "Recovered", "Guam", "Grand Princess",
                  "District of Columbia", "Diamond Princess")) %>% 
  filter(Lat > 0)

c <- ggplot(daily_report_04_12_2020, aes(x = Long, y = Lat, size = Confirmed)) +
    borders("state", colour = "white", fill = "grey90") +
    geom_point(aes(x=Long, y=Lat, size=Confirmed, color=Confirmed),stroke=F, alpha=0.7) +
    scale_size_continuous(name="Cases", trans="log", range=c(1,7), 
                        breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+")) +
    scale_color_viridis_c(option="viridis",name="Cases",
                        trans="log", breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+"))  +
# Cleaning up the graph
  
  theme_void() + 
    guides( colour = guide_legend()) +
    labs(title = "COVID-19 Confirmed Cases in the US by April 12, 2020") +
    theme(
      legend.position = "bottom",
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#ffffff", color = NA), 
      panel.background = element_rect(fill = "#ffffff", color = NA), 
      legend.background = element_rect(fill = "#ffffff", color = NA)
    ) +
    coord_fixed(ratio=1.5)
```
```{r c, echo=FALSE, warnings=FALSE, message=FALSE}
c
```

##Number of Confirmed Cases by US County April 12, 2020
```{r, warnings=FALSE, message=FALSE, include=FALSE, echo=FALSE}
report_04_12_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-12-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  unite(Key, Admin2, Province_State, sep = ".") %>% 
  group_by(Key) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Key = tolower(Key))
# dim(report_03_27_2020)
# get and format the map data
us <- map_data("state")
counties <- map_data("county") %>% 
  unite(Key, subregion, region, sep = ".", remove = FALSE)
# Join the 2 tibbles
state_join <- left_join(counties, report_04_12_2020, by = c("Key"))
# sum(is.na(state_join$Confirmed))

d <- ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  # Add data layer
  borders("state", colour = "black") +
  geom_polygon(data = state_join, aes(fill = Confirmed)) +
  scale_fill_gradientn(colors = brewer.pal(n = 5, name = "PuRd"),
                       breaks = c(1, 10, 100, 1000, 10000, 100000),
                       trans = "log10", na.value = "White") +
  ggtitle("Number of Confirmed Cases by US County by April 12, 2020") +
  theme_bw() 
```
```{r d, echo=FALSE, warnings=FALSE, message=FALSE}
d
```

##COVID-19 Confirmed Cases by county in Massachusetts April 12, 2020
```{r warnings=FALSE, message=FALSE, include=FALSE, echo=FALSE}
daily_report_04_12_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-12-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "Massachusetts") %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))
us <- map_data("state")
ma_us <- subset(us, region == "massachusetts")
counties <- map_data("county")
ma_county <- subset(counties, region == "massachusetts")
state_join <- left_join(ma_county, daily_report_04_12_2020, by = c("subregion" = "Admin2")) 
# plot state map
e <- ggplot(data = ma_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "white") +
    scale_fill_gradientn(colors = brewer.pal(n = 5, name = "BuGn"),
                         trans = "log10") +
  labs(title = "COVID-19 Confirmed Cases in Massachusetts by April 12, 2020")+
  theme_bw() 
```
```{r e, echo=FALSE, warnings=FALSE, message=FALSE}
e
```

##COVID-19 Confirmed Cases by county in Massachusetts April 12, 2020
```{r warnings=FALSE, message=FALSE, include=FALSE, echo=FALSE}
library(plotly)
  f <- ggplot(data = ma_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "black") +
    scale_fill_gradientn(colours = wes_palette("Zissou1", 100, type = "continuous")) + 
  ggtitle("COVID-19 Cases in MA")  +
# Cleaning up the graph
  labs(x=NULL, y=NULL) +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank())
```
```{r f, echo=FALSE, warnings=FALSE, message=FALSE}
  ggplotly(f)
```


##COVID-19 Confirmed Cases by county in New York April 12, 2020
```{r warnings=FALSE, message=FALSE, include=FALSE, echo=FALSE}
library(plotly)
daily_report_04_12_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-12-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "New York") %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))
us <- map_data("state")
NY_us <- subset(us, region == "new york")
counties <- map_data("county")
NY_county <- subset(counties, region == "new york")
state_join <- left_join(NY_county, daily_report_04_12_2020, by = c("subregion" = "Admin2")) 
  z <- ggplot(data = NY_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "black") +
    scale_fill_gradientn(colours = wes_palette("Zissou1", 100, type = "continuous")) + 
  ggtitle("COVID-19 Cases in New York")  +
# Cleaning up the graph
  labs(x=NULL, y=NULL) +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank())
```
```{r z, echo=FALSE, warnings=FALSE, message=FALSE}
  ggplotly(z)
```

##COVID-19 Confirmed Cases by county in California April 12, 2020
```{r warnings=FALSE, message=FALSE, include=FALSE, echo=FALSE}
library(plotly)
daily_report_04_12_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-12-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "California") %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))
us <- map_data("state")
ca_us <- subset(us, region == "california")
counties <- map_data("county")
ca_county <- subset(counties, region == "california")
state_join <- left_join(ca_county, daily_report_04_12_2020, by = c("subregion" = "Admin2")) 
  g <- ggplot(data = ca_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "black") +
    scale_fill_gradientn(colours = wes_palette("Zissou1", 100, type = "continuous")) + 
  ggtitle("COVID-19 Cases in California")  +
# Cleaning up the graph
  labs(x=NULL, y=NULL) +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank())
```
```{r g, echo=FALSE, warnings=FALSE, message=FALSE}
  ggplotly(g)
```


```{r warnings=FALSE, message=FALSE, include=FALSE, echo=FALSE}
#China confirmed cases
#daily_report_03_27_2020 %>% 
#  filter(Country_Region == "China") %>% 
#  filter(Lat > 0) %>% 
#  ggplot(aes(x = Long, y = Lat, size = Confirmed/1000)) +
#    borders("china", colour = "grey", fill = "grey90") +
#    theme_bw() +
#    geom_point(shape = 21, color='blue', fill='blue', alpha = 0.5) +
#    labs(title = 'COVID-19 Confirmed Cases in China by March 27, 2020', x = '', y = '',
#        size="Cases (x1000))") +
#    theme(legend.position = "right") +
#    coord_fixed(ratio=1.5)
```

```{r, warnings=FALSE, message=FALSE, include=FALSE}
#create data.frame for China map
chinamap <- map_data("world", region = "china")
region.coord.china <- chinamap %>% 
  group_by(subregion) %>%
  summarise(long = mean(long), lat = mean(lat))
```

```{r, warnings=FALSE, message=FALSE, include=FALSE}
#filter dataset to add names to map
chinamap2 <- daily_report_03_31_2020 %>% 
  filter(Country_Region == "China") %>% 
  select(Lat,Long,Province_State)

#region.coord.china2 <- chinamap2 %>% 
# summarise(long = mean(Long), lat = mean(Lat))
```

##COVID-19 Confirmed Cases in China March 31, 2020
```{r, warnings=FALSE, message=FALSE, include=FALSE, echo=TRUE}
mybreaks <- c(1, 100, 1000, 10000, 10000)
daily_report_03_31_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-31-2020.csv")) %>% 
  select(-Admin2, -FIPS) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "China") %>% 
  filter(Lat > 0) 
h <- daily_report_03_31_2020 %>% 
  filter(Country_Region == "China") %>% 
  filter(Lat > 0) %>% 
  ggplot()+
  geom_polygon(data = chinamap, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3,colour='black') +
    #borders("china", colour = "black", fill = "grey90") + don't include this!
    geom_point(aes(x=Long, y=Lat, size=Confirmed, color=Confirmed),stroke=F, alpha=0.7) +
    scale_size_continuous(name="Cases", trans="log", range=c(1,7), 
                        breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+")) +
    scale_color_viridis_c(option="viridis",name="Cases",
                        trans="log", breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+"))  +
# Cleaning up the graph
  
  theme_void() + 
    guides( colour = guide_legend()) +
    labs(title = "COVID-19 Confirmed Cases in China") +
    theme(
      legend.position = "bottom",
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#ffffff", color = NA), 
      panel.background = element_rect(fill = "#ffffff", color = NA), 
      legend.background = element_rect(fill = "#ffffff", color = NA)
    ) +
    coord_fixed(ratio=1.5)+
  geom_text(aes(x=Long, y=Lat,label=Province_State), data = chinamap2, size =3, hjust=0.5)
  
```
```{r h, echo=FALSE, warnings=FALSE, message=FALSE}
h
```

```{r warnings=FALSE, message=FALSE, include=FALSE, echo=FALSE}
###Confirmed
time_series_confirmed_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")  %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Confirmed") 

###Deaths
time_series_deaths_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")  %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Deaths")

###Recovered
time_series_recovered_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region") %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Recovered")

# Create Keys 
time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)
time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)
time_series_recovered_long <- time_series_recovered_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Recovered)
# Join tables
time_series_long_joined <- full_join(time_series_confirmed_long,
              time_series_deaths_long, by = c("Key"))
time_series_long_joined <- full_join(time_series_long_joined,
              time_series_recovered_long, by = c("Key")) %>% 
    select(-Key)
# Reformat the data
time_series_long_joined$Date <- mdy(time_series_long_joined$Date)
# Create Report table with counts
time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date),
               names_to = "Report_Type", values_to = "Counts")
```

```{r warnings=FALSE, message=FALSE, include=FALSE, echo=FALSE}
library(ggplot2)
library(gganimate)
library(transformr)
theme_set(theme_bw())
```


##Confirmed cases in china, Italy, Japan, South Korea and US
```{r, warnings=FALSE, message=FALSE, include=FALSE, echo=FALSE}
data_time <- time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths", "Recovered"), sum) %>% 
    filter (Country_Region %in% c("China","Korea, South","Japan","Italy","US")) 

p <- ggplot(data_time, aes(x = Date,  y = Confirmed, color = Country_Region)) + 
      geom_point() +
      geom_line() +
      ggtitle("Confirmed COVID-19 Cases") +
      geom_point(aes(group = seq_along(Date))) +
      transition_reveal(Date) 
# Some people needed to use this line instead
# animate(p,renderer = gifski_renderer(), end_pause = 15)
```

```{r p, warnings=FALSE, message=FALSE, echo=TRUE}
animate(p, end_pause = 15)
```

##Confirmed cases over time wordlwide
```{r, warnings=FALSE, message=FALSE, include=FALSE, echo=FALSE}
covid <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
           rename(Province_State= "Province/State", Country_Region = "Country/Region") %>%
           pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                  names_to = "Date", values_to = "Confirmed") %>%
           mutate(Date = mdy(Date) - days(1),
                  Place = paste(Lat,Long,sep="_")) %>%
# Summarizes state and province information
             group_by(Place,Date) %>%
           summarise(cumulative_cases = ifelse(sum(Confirmed)>0,
                     sum(Confirmed),NA_real_),
                     Lat = mean(Lat),
                     Long = mean(Long)) %>%
           mutate(Pandemic_day = as.numeric(Date - min(Date)))
world <- ggplot(covid,aes(x = Long, y = Lat, size = cumulative_cases/1000)) +
                 borders("world", colour = "gray50", fill = "grey90") +
                 theme_bw() +
                 geom_point(color='purple', alpha = .5) +
                 labs(title = 'Pandemic Day: {frame}',x = '', y = '',
                      size="Cases (x1000))") +
                 theme(legend.position = "right") +
                 coord_fixed(ratio=1.3)+
                 transition_time(Date) +
                 enter_fade()
# Some people needed to use this line instead
# animate(world,renderer = gifski_renderer(), end_pause = 30)

```
```{r world, warnings=FALSE, message=FALSE, echo=FALSE}
animate(world, end_pause = 10)
```

##Confirmed cases in Latin America 
```{r, warnings=FALSE, message=FALSE, include=FALSE, echo=FALSE}
data_time <- time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths", "Recovered"), sum) %>% 
    filter (Country_Region %in% c("Colombia","Brazil","Argentina","Chile","Ecuador", "Peru", "Venezuela")) 

l <- ggplot(data_time, aes(x = Date,  y = Confirmed, color = Country_Region)) + 
      geom_point() +
      geom_line() +
      ggtitle("Confirmed COVID-19 Cases in Latin America") +
      geom_point(aes(group = seq_along(Date))) +
      transition_reveal(Date) 
# Some people needed to use this line instead
# animate(p,renderer = gifski_renderer(), end_pause = 15)
```

```{r l, warnings=FALSE, message=FALSE, echo=FALSE}
animate(l, end_pause = 15)
```






