---
title: "Lab10 Covid19 maps"
author: "Katherine Chacon-Vargas"
date: "4/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warmings=FALSE, message=FALSE}
library(tidyverse)
library(maps)
library(mapdata)
library(lubridate)
library(viridis)
library(wesanderson)
library(RColorBrewer)
```


```{r, warnings=FALSE, message=FALSE}
daily_report_03_31_2020 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-31-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  select(-Admin2, -FIPS)
    
ggplot(daily_report_03_31_2020, aes(x = Long, y = Lat, size = Confirmed/1000)) +
    borders("world", colour = NA, fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='blue', fill='blue', alpha = 0.5) +
    labs(title = 'World COVID-19 Confirmed cases',x = '', y = '',
        size="Cases (x1000))") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```

#US states confirmed cases 
```{r, warnings=FALSE, message=FALSE}
daily_report_03_27_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-27-2020.csv")) %>% 
  select(-Admin2, -FIPS) %>% 
  rename(Long = "Long_") 
```

```{r, warnings=FALSE, message=FALSE}
daily_report_03_27_2020 %>% 
  filter(Country_Region == "US") %>% 
  filter (!Province_State %in% c("Alaska","Hawaii", "American Samoa",
                  "Puerto Rico","Northern Mariana Islands", 
                  "Virgin Islands", "Recovered", "Guam", "Grand Princess",
                  "District of Columbia", "Diamond Princess")) %>% 
  filter(Lat > 0) %>% 
  ggplot(aes(x = Long, y = Lat, size = Confirmed/1000)) +
    borders("state", colour = "black", fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='blue', fill='blue', alpha = 0.5) +
    labs(title = 'COVID-19 Confirmed Cases in the US by March 27, 2020', x = '', y = '',
        size="Cases (x1000))") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```
 

```{r, warnings=FALSE, message=FALSE}
#China confirmed cases
#daily_report_03_27_2020 %>% 
#  filter(Country_Region == "China") %>% 
#  filter(Lat > 0) %>% 
#  ggplot(aes(x = Long, y = Lat, size = Confirmed/1000)) +
#    borders("china", colour = "grey", fill = "grey90") +
#    theme_bw() +
#    geom_point(shape = 21, color='blue', fill='blue', alpha = 0.5) +
#    labs(title = 'COVID-19 Confirmed Cases in China by March 27, 2020', x = '', y = '',
#        size="Cases (x1000))") +
#    theme(legend.position = "right") +
#    coord_fixed(ratio=1.5)
```
 

##US
```{r, warnings=FALSE, message=FALSE}
mybreaksUS <- c(1, 100, 1000, 10000, 10000)
daily_report_03_27_2020 %>% 
  filter(Country_Region == "US") %>% 
  filter (!Province_State %in% c("Alaska","Hawaii", "American Samoa",
                  "Puerto Rico","Northern Mariana Islands", 
                  "Virgin Islands", "Recovered", "Guam", "Grand Princess",
                  "District of Columbia", "Diamond Princess")) %>% 
  filter(Lat > 0) %>% 
ggplot(aes(x = Long, y = Lat, size = Confirmed)) +
    borders("state", colour = "white", fill = "grey90") +
    geom_point(aes(x=Long, y=Lat, size=Confirmed, color=Confirmed),stroke=F, alpha=0.7) +
    scale_size_continuous(name="Cases", trans="log", range=c(1,7), 
                        breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+")) +
    scale_color_viridis_c(option="viridis",name="Cases",
                        trans="log", breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+"))  +
# Cleaning up the graph
  
  theme_void() + 
    guides( colour = guide_legend()) +
    labs(title = "COVID-19 Confirmed Cases in the US'") +
    theme(
      legend.position = "bottom",
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#ffffff", color = NA), 
      panel.background = element_rect(fill = "#ffffff", color = NA), 
      legend.background = element_rect(fill = "#ffffff", color = NA)
    ) +
    coord_fixed(ratio=1.5)
```


##China

```{r, warnings=FALSE, message=FALSE}
#create data.frame for China map
chinamap <- map_data("world", region = "china")
region.coord.china <- chinamap %>% 
  group_by(subregion) %>%
  summarise(long = mean(long), lat = mean(lat))
```

```{r, warnings=FALSE, message=FALSE}
#filter dataset to add names to map
chinamap2 <- daily_report_03_27_2020 %>% 
  filter(Country_Region == "China") %>% 
  select(Lat,Long,Province_State)

#region.coord.china2 <- chinamap2 %>% 
# summarise(long = mean(Long), lat = mean(Lat))
```

```{r, warnings=FALSE, message=FALSE}
mybreaks <- c(1, 100, 1000, 10000, 10000)
daily_report_03_27_2020 %>% 
  filter(Country_Region == "China") %>% 
  filter(Lat > 0) %>% 
  ggplot()+
  geom_polygon(data = chinamap, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3,colour='black') +
    #borders("china", colour = "black", fill = "grey90") + don't include this!
    geom_point(aes(x=Long, y=Lat, size=Confirmed, color=Confirmed),stroke=F, alpha=0.7) +
    scale_size_continuous(name="Cases", trans="log", range=c(1,7), 
                        breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+")) +
    scale_color_viridis_c(option="viridis",name="Cases",
                        trans="log", breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+"))  +
# Cleaning up the graph
  
  theme_void() + 
    guides( colour = guide_legend()) +
    labs(title = "COVID-19 Confirmed Cases in China") +
    theme(
      legend.position = "bottom",
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#ffffff", color = NA), 
      panel.background = element_rect(fill = "#ffffff", color = NA), 
      legend.background = element_rect(fill = "#ffffff", color = NA)
    ) +
    coord_fixed(ratio=1.5)+
  geom_text(aes(x=Long, y=Lat,label=Province_State), data = chinamap2, size =3, hjust=0.5)
  
```

* Note that in both examples the ggplot funtion __borders__ is used to define the areas in the map

### Mapping data to shapes

```{r, warnings=FALSE, message=FALSE}
daily_report <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-02-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  group_by(Province_State) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Province_State = tolower(Province_State))
# load the US map data
us <- map_data("state")
# We need to join the us map data with our daily report to make one data frame/tibble
state_join <- left_join(us, daily_report, by = c("region" = "Province_State"))
# plot state map
ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "white") +
  geom_polygon(color = "black", fill = NA) +
  scale_fill_gradient(trans = "log10") +
  labs(title = "COVID-19 Confirmed Cases in the US'")
```

#### Using R color palattes

This is a bit of a digression back to Labs 3 and 4, but there are many R color palattes to choose from or you can create your own. In the above a simple gradient is used. The example from Anisa Dhana uses the viridis palatte which is designed to be perceived by viewers with common forms of colour blindness. Here is an example using a different color package - Wes Anderson. ...and more [	
Top R Color Palettes to Know for Great Data Visualization](https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/)

```{r, warnings=FALSE, message=FALSE}
# plot state map
ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "white") +
  geom_polygon(color = "black", fill = NA) +
  scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous"),
                         trans = "log10") +
  labs(title = "COVID-19 Confirmed Cases in the US'")
```

If we look at just Massachusetts

```{r, warnings=FALSE, message=FALSE}
daily_report_04_02_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-02-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "Massachusetts") %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))
us <- map_data("state")
ma_us <- subset(us, region == "massachusetts")
counties <- map_data("county")
ma_county <- subset(counties, region == "massachusetts")
state_join <- left_join(ma_county, daily_report, by = c("subregion" = "Admin2")) 
# plot state map
ggplot(data = ma_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "white") +
    scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous"),
                         trans = "log10") +
  labs(title = "COVID-19 Confirmed Cases in Massachusetts'")
```

* Note the cases on Nantucket and Dukes counties were reported as one value and not included on the graph. There is also an asssigned category that includes 303 Confirmed cases as of 3/31/2020.
```{r}
daily_report
```

## Interactive graphs

In Lab 4 plotly was introduced. It is a great simple way to make interactive graphs

```{r, warnings=FALSE, message=FALSE}
library(plotly)
ggplotly(
  ggplot(data = ma_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "black") +
    scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous")) +
  ggtitle("COVID-19 Cases in MA") +
# Cleaning up the graph
  labs(x=NULL, y=NULL) +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank())
)
```

## Animated Graphs with gganimate

Animated graphs when down right have a great visual impact. You can do this in R and have your animations embedded on your web page. Essentially gganimate creates a series of files that are encompassed in a gif file. In addition to having this gif as part of your report file, you can save the gif and use in a slide or other presentations. It just takes a few lines of code to covert and existing ggplot graph into an animation. See [Tutorial for Getting Started with gganimate](https://gganimate.com/articles/gganimate.html) and
[gganimate: How to Create Plots with Beautiful Animation in R](https://www.datanovia.com/en/blog/gganimate-how-to-create-plots-with-beautiful-animation-in-r/)

This are some important gganimate functions:

* __transition_*()__ defines how the data should be spread out and how it relates to itself across time.
* __view_*()__ defines how the positional scales should change along the animation.
* __shadow_*()__ defines how data from other points in time should be presented in the given point in time.
*  __enter_*()/exit_*()__ defines how new data should appear and how old data should disappear during the course of the animation.
* __ease_aes()__ defines how different aesthetics should be eased during transitions.

### Preparing the times series data

The time series data is ripe for animation but first we need to get and format the files

```{r, warnings=FALSE, message=FALSE}
time_series_confirmed_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")  %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Confirmed") 
# Let's get the times series data for deaths
time_series_deaths_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")  %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Deaths")
time_series_recovered_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region") %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Recovered")
# Create Keys 
time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)
time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)
time_series_recovered_long <- time_series_recovered_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Recovered)
# Join tables
time_series_long_joined <- full_join(time_series_confirmed_long,
              time_series_deaths_long, by = c("Key"))
time_series_long_joined <- full_join(time_series_long_joined,
              time_series_recovered_long, by = c("Key")) %>% 
    select(-Key)
# Reformat the data
time_series_long_joined$Date <- mdy(time_series_long_joined$Date)
# Create Report table with counts
time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date),
               names_to = "Report_Type", values_to = "Counts")
```

### Creating the animations

Below are the packages I installed. There may be others that you need, in particular to rendering gifs. Some of the example may take several minutes to create the animation.

```{r, warnings=FALSE, message=FALSE}
library(ggplot2)
library(gganimate)
library(transformr)
theme_set(theme_bw())
```

```{r}
install.packages("sf",dependencies = T)
install.packages("devtools")
devtools::install_github("thomasp85/transformr")
```


An animation of the confirmed cases in select countries

```{r, warnings=FALSE, message=FALSE}
data_time <- time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths", "Recovered"), sum) %>% 
    filter (Country_Region %in% c("China","Korea, South","Japan","US")) 
p <- ggplot(data_time, aes(x = Date,  y = Confirmed, color = Country_Region)) + 
      geom_point() +
      geom_line() +
      ggtitle("Confirmed COVID-19 Cases") +
      geom_point(aes(group = seq_along(Date))) +
      transition_reveal(Date) 
    
animate(p, end_pause = 15)
```

Animation for Prof. Chris Sunderland's example

```{r, warnings=FALSE, message=FALSE}
covid <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
           rename(Province_State= "Province/State", Country_Region = "Country/Region") %>%
           pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                  names_to = "Date", values_to = "Confirmed") %>%
           mutate(Date = mdy(Date) - days(1),
                  Place = paste(Lat,Long,sep="_")) %>%
# Summarizes state and province information
             group_by(Place,Date) %>%
           summarise(cumulative_cases = ifelse(sum(Confirmed)>0,
                     sum(Confirmed),NA_real_),
                     Lat = mean(Lat),
                     Long = mean(Long)) %>%
           mutate(Pandemic_day = as.numeric(Date - min(Date)))
world <- ggplot(covid,aes(x = Long, y = Lat, size = cumulative_cases/1000)) +
                 borders("world", colour = "gray50", fill = "grey90") +
                 theme_bw() +
                 geom_point(color='purple', alpha = .5) +
                 labs(title = 'Pandemic Day: {frame}',x = '', y = '',
                      size="Cases (x1000))") +
                 theme(legend.position = "right") +
                 coord_fixed(ratio=1.3)+
                 transition_time(Date) +
                 enter_fade()
animate(world, end_pause = 30)
```
