---
title: "Lab 9 Covid-19"
author: "Katherine Chacon-Vargas"
date: "3/25/2020"
output: html_document
---

```{r message=FALSE}
library(tidyverse)
library(lubridate)
library(RColorBrewer)
```

```{r}
report_03_11_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-11-2020.csv")) %>%
  rename(Country.Region = "Country/Region", Province.State = "Province/State")
head(report_03_11_2020)
str(report_03_11_2020)
```

```{r}
report_03_24_2020 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-24-2020.csv")) %>% 
  rename(Country.Region = "Country_Region", Province.State = "Province_State") %>% 
  select(-FIPS, -Admin2)
head(report_03_24_2020)
str(report_03_24_2020)
```

##Cases in US states as of March 11th
```{r}
report_03_11_2020 %>% 
  filter(Country.Region =="US") %>% 
  ggplot(aes(Confirmed,reorder(Province.State,Confirmed)))+
  geom_point() +
  ggtitle("Confirmed cases for each US State")+
  ylab("Country/Region")+
  xlab("Confirmed Cases")

ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/Cases_US_march_11.png", width = 10, height = 10, dpi = "screen")
```
```{r}
report_03_11_2020 %>% 
  group_by(Country.Region) %>% 
  summarise(Deaths=sum(Deaths)) %>% 
  arrange(desc(Deaths))
```
```{r}
report_03_11_2020 %>% 
  group_by(Country.Region) %>% 
  summarise(Deaths=sum(Deaths)) %>% 
  arrange(desc(Deaths)) %>% 
  slice(1:20) %>% 
  ggplot(aes(x=Deaths, y=reorder(Country.Region, Deaths)))+
  geom_bar(stat='identity')+
    ggtitle("The 20 countries with the most reported COV19-related deaths") +
  ylab("Country/Region")+
  xlab("Deaths")

ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/Deaths_20_countries_march_11.png", width = 10, height = 10, dpi = "screen")
```

###Times Series Data
```{r}
time_series_confirmed <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province.State = "Province/State", Country.Region = "Country/Region")

head(time_series_confirmed)
str(time_series_confirmed)

```
```{r}
time_series_deaths <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")) %>%
  rename(Province.State = "Province/State", Country.Region = "Country/Region")

head(time_series_deaths)
str(time_series_deaths)

```

```{r}
time_series_confirmed_long <- time_series_confirmed %>% 
               pivot_longer(-c(Province.State, Country.Region, Lat, Long),
                            names_to = "Date", values_to = "Confirmed") %>% 
               group_by(Country.Region,Date) %>% 
               summarise(Confirmed = sum(Confirmed))

# convert date to data format
time_series_confirmed_long$Date <- mdy(time_series_confirmed_long$Date)
head(time_series_confirmed_long)
```


#US confirmed COVID-19 Cases
```{r}
time_series_confirmed_long %>% 
  filter (Country.Region == "US") %>% 
    ggplot(aes(x = Date,  y = Confirmed)) + 
    geom_point() +
    geom_line() +
    ggtitle("US Confirmed COVID-19 Cases")

ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/US_all_march_11.png", width = 10, height = 10, dpi = "screen")
```

#Colombia confirmed COVID-19 Cases
```{r}
time_series_confirmed_long %>% 
  filter (Country.Region == "Colombia") %>% 
    ggplot(aes(x = Date,  y = Confirmed)) + 
    geom_point() +
    geom_line() +
    ggtitle("Colombia Confirmed COVID-19 Cases")

ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/Cases_colombia_march_11.png", width = 10, height = 10, dpi = "screen")
```

#China confirmed COVID-19 Cases
```{r}
time_series_confirmed_long %>% 
  filter (Country.Region == "China") %>% 
    ggplot(aes(x = Date,  y = Confirmed)) + 
    geom_point() +
    geom_line() +
    ggtitle("China Confirmed COVID-19 Cases")

ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/Cases_china_march_11.png", width = 10, height = 10, dpi = "screen")
```

```{r}
time_series_confirmed_long %>% 
  filter(Country.Region %in% c("China", "Japan", "Korea, South","Italy","Spain","US" )) %>% 
  ggplot(aes(x=Date, y=Confirmed))+
  geom_point()+
  geom_line()+
  ggtitle("Confirmed COVID-19 Cases") +
  facet_wrap(~Country.Region, ncol = 2, scales = "free_y")

ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/Cases_comparison_march_11.png", width = 10, height = 10, dpi = "screen")

```

##Latin America COVID-19 Cases
```{r}
time_series_confirmed_long %>% 
  filter(Country.Region %in% c("Colombia", "Chile", "Argentina","Brazil","Ecuador","Peru","Mexico","Venezuela" )) %>% 
  ggplot(aes(x=Date, y=Confirmed))+
  geom_point()+
  geom_line()+
  ggtitle("Confirmed COVID-19 Cases") +
  facet_wrap(~Country.Region, ncol = 2, scales = "free_y")

ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/Cases_LatinAmerica_comparison_march_11.png", width = 10, height = 10, dpi = "screen")

```

##Same graph 
```{r}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
time_series_confirmed_long %>% 
  filter (Country.Region %in% c("China","France","Italy", 
                                "Korea, South", "US")) %>% 
    ggplot(aes(x = Date,  y = Confirmed, color = Country.Region)) + 
    geom_point() +
    scale_color_manual(values = cbPalette)+
    geom_line() +
    ggtitle("Confirmed COVID-19 Cases")
ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/Cases_all_march_11.png", width = 10, height = 10, dpi = "screen")
```

```{r}
time_series_confirmed_long %>% 
  filter (Country.Region %in% c("Colombia", "Chile", "Argentina","Brazil","Ecuador","Peru","Mexico","Venezuela")) %>% 
    ggplot(aes(x = Date,  y = Confirmed, color = Country.Region)) + 
    geom_point() +
    geom_line() +
   scale_color_manual(values = cbPalette)+
    ggtitle("Confirmed COVID-19 Cases")
ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/Cases_all2_march_11.png", width = 10, height = 10, dpi = "screen")
```

```{r}
report_03_24_2020 %>% 
    filter (Country.Region == "US") %>% 
    group_by(Province.State) %>% 
    summarise(Confirmed = sum(Confirmed)) %>% 
    ggplot(aes(x = Confirmed, y = reorder(Province.State, Confirmed))) + 
    geom_point() + 
    ggtitle("Confirmed cases for each US State") + 
    ylab ("Country/Region") + 
    xlab("Confirmed Cases March 24") 

ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/Cases_US_march_24.png", width = 10, height = 10, dpi = "screen")
```

```{r}
report_03_24_2020 %>% 
    group_by(Country.Region) %>% 
    summarise(Deaths = sum(Deaths)) %>% 
    arrange(desc(Deaths)) %>% 
    slice(1:20) %>% 
    ggplot(aes(y = Deaths, x = reorder(Country.Region, Deaths))) + 
    geom_bar(stat = 'identity') + 
    ggtitle("The 20 Countries with the most reported COV19-related deaths March 24")+
    ylab("Country/Region") + 
    xlab("Deaths") + 
    coord_flip() 

ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/deaths_world_march_24.png", width = 10, height = 10, dpi = "screen")
```

```{r}
report_03_24_2020 %>% 
    group_by(Country.Region) %>% 
    summarise(Deaths = sum(Deaths)) %>% 
    arrange(desc(Deaths)) %>% 
    slice(1:20) %>% 
    ggplot(aes(y = Deaths, x = reorder(Country.Region, Deaths))) + 
    geom_bar(stat = 'identity') + 
    ggtitle("The 20 Countries with the most reported COV19-related deaths")+
    ylab("Country/Region") + 
    xlab("Deaths") + 
    coord_flip() 
```

```{r message=FALSE}
library(maps)
library(viridis)
```

```{r message=FALSE}
world <- map_data("world")
```

```{r message=FALSE}
mybreaks <- c(1, 20, 100, 1000, 50000)
```

##Confirmed cases 2/25/20
```{r message=FALSE}
ggplot() +
  geom_polygon(data = world, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3) +
  geom_point(data=time_series_confirmed, aes(x=Long, y=Lat, size=`2/25/20`, color=`2/25/20`),stroke=F, alpha=0.7) +
  scale_size_continuous(name="Cases", trans="log", range=c(1,7),breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000+")) +
  # scale_alpha_continuous(name="Cases", trans="log", range=c(0.1, 0.9),breaks=mybreaks) +
  scale_color_viridis_c(option="inferno",name="Cases", trans="log",breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000+")) +
  theme_void() + 
  guides( colour = guide_legend()) +
  labs(caption = "") +
  theme(
    legend.position = "bottom",
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#ffffff", color = NA), 
    panel.background = element_rect(fill = "#ffffff", color = NA), 
    legend.background = element_rect(fill = "#ffffff", color = NA)
  )+
      ggtitle("Confirmed COVID-19 Cases Feb 25/20")
ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/onfirmed_world_feb_25.png", dpi = "screen")
```
##Covid-19 deaths 2/25/20

```{r message=FALSE}
ggplot() +
  geom_polygon(data = world, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3) +
  geom_point(data=time_series_deaths, aes(x=Long, y=Lat, size=`2/25/20`, color=`2/25/20`),stroke=F, alpha=0.7) +
  scale_size_continuous(name="Cases", trans="log", range=c(1,7),breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000+")) +
  # scale_alpha_continuous(name="Cases", trans="log", range=c(0.1, 0.9),breaks=mybreaks) +
  scale_color_viridis_c(option="inferno",name="Cases", trans="log",breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000+")) +
  theme_void() + 
  guides( colour = guide_legend()) +
  labs(caption = "") +
  theme(
    legend.position = "bottom",
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#ffffff", color = NA), 
    panel.background = element_rect(fill = "#ffffff", color = NA), 
    legend.background = element_rect(fill = "#ffffff", color = NA)
  )+
  ggtitle("Confirmed COVID-19 deaths Feb 25/20")
ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/deaths_world_feb_25.png", dpi = "screen")

```

##Confirmed cases 3/24/20
```{r message=FALSE}
ggplot() +
  geom_polygon(data = world, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3) +
  geom_point(data=time_series_confirmed, aes(x=Long, y=Lat, size=`3/24/20`, color=`3/24/20`),stroke=F, alpha=0.7) +
  scale_size_continuous(name="Cases", trans="log", range=c(1,7),breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000+")) +
  # scale_alpha_continuous(name="Cases", trans="log", range=c(0.1, 0.9),breaks=mybreaks) +
  scale_color_viridis_c(option="inferno",name="Cases", trans="log",breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000+")) +
  theme_void() + 
  guides( colour = guide_legend()) +
  labs(caption = "") +
  theme(
    legend.position = "bottom",
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#ffffff", color = NA), 
    panel.background = element_rect(fill = "#ffffff", color = NA), 
    legend.background = element_rect(fill = "#ffffff", color = NA)
  )+
  ggtitle("Confirmed COVID-19 Cases March 24/20")

ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/confirmed_world_march_24.png", dpi = "screen")
```


##Covid-19 deaths 3/24/20
```{r message=FALSE}
ggplot() +
  geom_polygon(data = world, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3) +
  geom_point(data=time_series_deaths, aes(x=Long, y=Lat, size=`3/24/20`, color=`3/24/20`),stroke=F, alpha=0.7) +
  scale_size_continuous(name="Deaths", trans="log", range=c(1,7),breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000+")) +
  # scale_alpha_continuous(name="Deaths", trans="log", range=c(0.1, 0.9),breaks=mybreaks) +
  scale_color_viridis_c(option="inferno",name="Deaths", trans="log",breaks=mybreaks, labels = c("1-19", "20-99", "100-999", "1,000-49,999", "50,000+")) +
  theme_void() + 
  guides( colour = guide_legend()) +
  labs(caption = "") +
  theme(
    legend.position = "bottom",
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#ffffff", color = NA), 
    panel.background = element_rect(fill = "#ffffff", color = NA), 
    legend.background = element_rect(fill = "#ffffff", color = NA)
  )+
   ggtitle("Confirmed COVID-19 deaths March 24/20")

ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/deaths_world_march_24.png", dpi = "screen")
```

```{r}
time_series_confirmed_long2 <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
    rename(Province.State = "Province/State", Country.Region = "Country/Region") %>%
  pivot_longer(-c(Province.State, Country.Region, Lat, Long),
    names_to = "Date", values_to = "cumulative_cases") %>%
    mutate(Date = mdy(Date) - days(1),
        Place = paste(Lat,Long,sep="_")) %>%
    group_by(Place,Date) %>%
        summarise(cumulative_cases = ifelse(sum(cumulative_cases)>0,
        sum(cumulative_cases),NA_real_),
        Lat = mean(Lat),
        Long = mean(Long)) %>%
    mutate(Pandemic_day = as.numeric(Date - min(Date)))

```

```{r}
head(time_series_confirmed_long2)
```

```{r}
 ggplot(subset(time_series_confirmed_long2, Date %in% seq(min(Date),max(Date),7)),
            aes(x = Long, y = Lat, size = cumulative_cases/1000)) +
            borders("world", colour = NA, fill = "grey90") +
            theme_bw() +
            geom_point(shape = 21, color='purple', fill='purple', alpha = 0.5) +
            labs(title = 'COVID-19 spread',x = '', y = '',
                 size="Cases (x1000))") +
            theme(legend.position = "right") +
            coord_fixed(ratio=1)+
            facet_wrap(.~Date,nrow=3)


ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/static.png", dpi = "screen")
```

##Latin America
```{r}
#Latin countries
some.latin <- c("Colombia","Brazil","Peru","Ecuador","Chile","Venezuela","Bolivia","Argentina","Paraguay","Uruguay")
#retrieve map data
some.latin <- map_data("world", region = some.latin)
```

```{r}
#Coordinates of countries
region.coord.data <- some.latin %>%
  group_by(region) %>%
  summarise(long = mean(long), lat = mean(lat))
#Confirmed cases of COVID-19 in Latin America 
time_series_confirmed_latin <- time_series_confirmed %>% 
  filter (Country.Region %in% c("Colombia","Brazil","Peru","Ecuador","Chile","Venezuela","Bolivia","Argentina","Paraguay","Uruguay"))
#Confirmed Deaths of COVID-19 in Latin America
time_series_deaths_latin <- time_series_deaths %>% 
  filter(Country.Region %in% c("Colombia","Brazil","Peru","Ecuador","Chile","Venezuela","Bolivia","Argentina","Paraguay","Uruguay"))
```

##COVID-19 Confirmed Cases in South America reported on March 24/20
```{r fig.width=14, fig.height=12}
mybreaks2 <- c(1, 50, 100, 200, 300, 1000, 2000)
ggplot() +
    geom_polygon(data = some.latin, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3,colour='white') +
  geom_point(data=time_series_confirmed_latin, aes(x=Long, y=Lat, size=`3/24/20`, color=`3/24/20`),stroke=F, alpha=0.7)+
  scale_size_continuous(name="Cases", trans="log", range=c(1,20),breaks=mybreaks2, labels = c("1-49", "50-99", "100-199", "200-299", "300-399", "1000-1999",'2000+')) +
  # scale_alpha_continuous(name="Cases", trans="log", range=c(0.1, 0.9),breaks=mybreaks) +
  scale_color_viridis_c(option="inferno",name="Cases", trans="log",breaks=mybreaks2, labels = c("1-49","50-99", "100-199", "200-299", "300-399", "1000-1999",'2000+')) +
  theme_void() + 
  guides( colour = guide_legend()) +
  labs(caption = "") +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 30),
    legend.title = element_text(size = 30),
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#ffffff", color = NA), 
    panel.background = element_rect(fill = "#ffffff", color = NA), 
    legend.background = element_rect(fill = "#ffffff", color = NA))+
  geom_text(aes(x=long, y=lat,label=region), data = region.coord.data, size =5, hjust=0.5)+
   ggtitle("COVID-19 Confirmed Cases in South America reported on March 24/20")

ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/COVID-19-Confirmed-Cases-South-America-March-24.png", width = 14, height = 12, dpi = "screen")
```

##COVID-19 Confirmed Deaths in South America reported on March 24/20
```{r fig.width=14, fig.height=12}
mybreaks3 <- c(5,15, 25, 35, 45)
ggplot() +
    geom_polygon(data = some.latin, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3,colour='white') +
  geom_point(data=time_series_deaths_latin, aes(x=Long, y=Lat, size=`3/24/20`, color=`3/24/20`),stroke=F, alpha=0.7)+
  scale_size_continuous(name="Deaths", trans="log", range=c(1,20),breaks=mybreaks3, labels = c('0-9', '10-19', '20-39', '40-49','50+')) +
  # scale_alpha_continuous(name="Deaths", trans="log", range=c(0.1, 0.9),breaks=mybreaks) +
  scale_color_viridis_c(option="inferno",name="Deaths", trans="log",breaks=mybreaks3, labels = c('0-9', '10-19', '20-39', '40-49', '50+')) +
  theme_void() + 
  guides( colour = guide_legend()) +
  labs(caption = "") +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 30),
    legend.title = element_text(size = 30),
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#ffffff", color = NA), 
    panel.background = element_rect(fill = "#ffffff", color = NA), 
    legend.background = element_rect(fill = "#ffffff", color = NA))+
  geom_text(aes(x=long, y=lat,label=region), data = region.coord.data, size =5, hjust=0.5)+
  ggtitle("COVID-19 Confirmed Deaths in South America reported on March 24/20")

ggsave("/Users/katherinechaconvargas/Dropbox/2020_UMASS/Human_Genome_Class_Jeffrey/images/COVID-19-Confirmed-deaths-South-America-March-24.png", width = 14, height = 12, dpi = "screen")
```


#Latin America
```{r}
#Latin cases 
time_series_confirmed_long2_latin <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
    rename(Province.State = "Province/State", Country.Region = "Country/Region") %>%
  filter (Country.Region %in% c("Colombia","Brazil","Peru","Ecuador","Chile","Venezuela","Bolivia","Argentina","Paraguay","Uruguay")) %>% 
  pivot_longer(-c(Province.State, Country.Region, Lat, Long),
    names_to = "Date", values_to = "cumulative_cases") %>%
    mutate(Date = mdy(Date) - days(1),
        Place = paste(Lat,Long,sep="_")) %>%
    group_by(Place,Date) %>%
        summarise(cumulative_cases = ifelse(sum(cumulative_cases)>0,
        sum(cumulative_cases),NA_real_),
        Lat = mean(Lat),
        Long = mean(Long)) %>%
    mutate(Pandemic_day = as.numeric(Date - min(Date))) %>% 
  filter(!is.na(cumulative_cases)) 
```
```{r fig.width=14, fig.height=12}
ggplot(subset(time_series_confirmed_long2_latin, Date %in% seq(min(Date),max(Date),4)),
            aes(x = Long, y = Lat, size = cumulative_cases/100))+
            borders('world',colour = NA, fill = "grey90", regions = c("Colombia","Brazil","Peru","Ecuador","Chile","Venezuela","Bolivia","Argentina","Paraguay","Uruguay"))+
  theme_bw() +
            geom_point(shape = 21, color='purple', fill='purple', alpha = 0.5) +
  labs(title = 'COVID-19 spread',x = '', y = '',
                 size="Cases (x1000))") +
            theme(legend.position = "right") +
            coord_fixed(ratio=1)+
            facet_wrap(.~Date,nrow=3)+
  geom_text(aes(x=long, y=lat,label=region), data = region.coord.data, size =2, hjust=0.5)

```


```{r}
map_colombia <- world %>% 
  filter(region == 'Colombia')
```
##COVID-19 spread Colombia
```{r}
#Colombia cases 
time_series_confirmed_long2_Colombia <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
    rename(Province.State = "Province/State", Country.Region = "Country/Region") %>%
  filter(Country.Region == 'Colombia') %>%
  pivot_longer(-c(Province.State, Country.Region, Lat, Long),
    names_to = "Date", values_to = "cumulative_cases") %>%
    mutate(Date = mdy(Date) - days(1),
        Place = paste(Lat,Long,sep="_")) %>%
    group_by(Place,Date) %>%
        summarise(cumulative_cases = ifelse(sum(cumulative_cases)>0,
        sum(cumulative_cases),NA_real_),
        Lat = mean(Lat),
        Long = mean(Long)) %>%
    mutate(Pandemic_day = as.numeric(Date - min(Date))) %>% 
filter(!is.na(cumulative_cases)) 
```

```{r fig.width=14, fig.height=12}
ggplot(subset(time_series_confirmed_long2_Colombia, Date %in% seq(min(Date),max(Date),3)),
            aes(x = Long, y = Lat, size = cumulative_cases/100))+
            borders('world',colour = NA, fill = "grey90", regions = 'Colombia')+
  theme_bw() +
            geom_point(shape = 21, color='purple', fill='purple', alpha = 0.5) +
  labs(title = 'COVID-19 spread',x = '', y = '',
                 size="Cases (x1000))") +
            theme(legend.position = "right") +
            coord_fixed(ratio=1)+
            facet_wrap(.~Date,nrow=3)

```


```